---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ints

<!-- badges: start -->
<!-- badges: end -->

[`{ints}`](ds-turner.github.io/ints/) helps to tidy up messy intervals. It can group and merge overlapping intervals as well as remove parts of intervals, based on other intervals. 

`{ints}` uses [`{data.table}`](https://rdatatable.gitlab.io/data.table/) for memory efficiency and speed, it will always return a `data.table` object. 

## Installation

You can install the development version of `{ints}` from [GitHub](https://github.com/) with:

``` r
pak::pak("ds-turner/ints")
```

## Useage
```{r}
library(ints)
```
### Grouping and Merging Intervals

We can group overlapping intervals by using `grp_ints`.  It returns the same data but add an id for intervlase that overlap
```{r}
 ints <- data.frame(
    id = c("A", "A", "A", "B", "B", "B"),
    st = c(1, 2, 5, 10, 12, 14),
    end = c(3, 4, 7, 11, 14, 16)
  )
 ints
```

```{r}
grp_ints(ints, st, end, id)
```
We can also merge overlapping intervals together using `merge_ints`.  It will return the interval start and end, the interval group id and any grouping variables used.
```{r}
merge_ints(ints, st, end, id)
```
### Triming Intervals

We can remove parts of intervals that overlap with intervals in another data frame.

Below is an example of some treatment data.  Each patient has a period monthly treatment for the duration of 2023.  For our analysis we will consider the different dose frequencies as mutually exclusive and hierarchical, i.e. a patient will not be considered as being dosed weekly if there is a record of daily dosing and a patient will not be considered as being dosed monthly if there is a record of weekly dosing.  
```{r}
trt <- data.frame(
  pat_id = c(1, 1, 1,
             2, 2, 2
             ),
  trt_start_date = as.Date(c(
    "2023-07-13", "2023-06-05", "2023-01-01",
    "2023-03-02", "2023-03-05", "2023-01-01"
  )),
  trt_end_date = as.Date(c(
    "2023-07-20", "2023-08-15", "2023-12-31",
    "2023-03-10", "2023-03-15", "2023-12-31"
  )),
  dose_freq = c(
    "daily", "weekly",  "monthly",
    "daily", "weekly", "monthly"
  )
)

trt
```

Here we split up the intervals to make sure everything is mutually exclusive.
```{r}
# create a list contaiing each dose level
trt_list <- split(trt, trt$dose_freq)

# remove the parts of the week dosing periods that overlap with the daily dosing intervals
trt_list$weekly <- trm_ints(
  trt_list$weekly,
  trt_list$daily,
  trt_start_date,
  trt_end_date,
  trt_start_date,
  trt_end_date,
  pat_id,
  .gap = 1
  )
# remove the parts of the monthly dosing periods that overlap with the daily and weekly dosing intervals
trt_list$monthly <- trm_ints(
  trt_list$monthly,
  rbind(trt_list$daily, trt_list$weekly),
  trt_start_date,
  trt_end_date,
  trt_start_date,
  trt_end_date,
  pat_id,
  .gap = 1
  )

trt2 <- data.table::rbindlist(
  trt_list
)
data.table::setorder(trt2, pat_id, trt_start_date)

trt2
```

### Negative Intervals
We can also create a set of intervals that represent the gaps between the given intervals
```{r}
ints <- data.frame(
  id = c("a", "a", "a", "a", "a", "b", "b", "c",   "c", "c", "c", "d"),
  start = c(1, 4, 10, 18, 23, 7, 12, 1, 7, 12, 23, 10),
  end = c(3, 7, 15, 21, 25, 9, 16, 3, 9, 16, 25, 15),
  index = c(5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5),
  study_end = c(20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20)
)
ints
```

```{r}
neg_ints(ints, start, end, id)
```
These intervals can have an upper and lower bounds if required
```{r}
neg_ints(ints, start, end, id, .lower = index, .upper = study_end)
```

#### Keywords
interval packing; interval merging; interval cutting;
