---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache =TRUE
)
```

# intervals

<!-- badges: start -->
<!-- badges: end -->

`{intervals}` helps users tidy up messy intervals. It can group and merge overlapping intervals as well as remove parts of one set of intervals, based on another set of intervals. In this context an interval is defined as having a start and a stop.  It can be numeric or date type.

{intervals} is mostly written in [`{data.table}`](https://rdatatable.gitlab.io/data.table/) for memory efficiency, it will always return a `data.table` object. 

## Installation

You can install the development version of intervals from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ds-turner/intervals")
```

## Useage
```{r}
library(intervals)
# library(data.table)
```
### Grouping and Packing Intervals
```{r}

 ints <- data.frame(
    id = c("A", "A", "A", "B", "B", "B"),
    st = c(1, 2, 5, 10, 12, 14),
    end = c(3, 4, 7, 11, 14, 16)
  )
 ints
```

`ints` is a set of overlapping intrevals.   We can identify the overlapping intervals and add an ID column `int_grp_id` for overlapping groups.
```{r}
grp_ints(ints, st, end, id)
```
We can also merge the overlapping intervels into single intervals.
```{r}
pac_ints(ints, st, end, id)
```
### Triming Intervals
You can remove parts of intervals based on another set of intervals.
```{r}
pats <- 7

x <- data.frame(
  id = c(1:pats),
  start = rep(5, pats),
  end = rep(20, pats)
)

y <- data.frame(
  id = c(1, 2, 3, 4, 5, 5, 5, 6, 6, 6, 6, 7, 7, 7, 7),
  start = c(10, 4, 4, 19, 4, 10, 19, 4, 9, 14, 19, 7, 9, 11, 13),
  end = c(15, 21, 6, 21, 6, 15, 21, 6, 11, 16, 21, 8, 10, 12, 18)
)
```

Here we will remove the parts of the intervals in x that overlap with the intervals in y.
```{r}
trm_ints(x, y, start, end, start, end, id)
```

### Negative Intervals
```{r}
ints <- data.frame(
  id = c("a", "a", "a", "a", "a", "b", "b", "c",   "c", "c", "c", "d"),
  start = c(1, 4, 10, 18, 23, 7, 12, 1, 7, 12, 23, 10),
  end = c(3, 7, 15, 21, 25, 9, 16, 3, 9, 16, 25, 15),
  index = c(5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5),
  study_end = c(20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20)
)
ints
```
We can also create a set of intervals that represent the gaps between the given intervals.
```{r}
neg_ints(ints, start, end, id)
```
These intervals can have an upper and lower bounds if required
```{r}
neg_ints(ints, start, end, id, .lower = index, .upper = study_end)
```
###
We have been given the below treatment data.  Each patient has a period monthly treatment for the duration of 2023.  For our analysis we will consider the different dose frequencies as mutually exclusive and higharical, i.e. a patient will not be considered as being dosed weekly if there is a record of daily dosing and a patient will not be considered as being dosed monthly if there is a record of weekly dosing.  
```{r}
trt <- data.frame(
  pat_id = c(1, 1, 1, 1,
             2, 2, 2
             ),
  trt_start_date = as.Date(c(
    "2023-07-13", "2023-06-05", "2023-07-01", "2023-01-01",
    "2023-03-02", "2023-03-05", "2023-01-01"
  )),
  trt_end_date = as.Date(c(
    "2023-07-20", "2023-08-15", "2023-10-13", "2023-12-31",
    "2023-03-10", "2023-03-15", "2023-12-31"
  )),
  dose_freq = c(
    "daily", "weekly",  "monthly", "monthly",
    "daily", "weekly", "monthly"
  )
)

trt
```
we can see that there is some overlapping so first we will merge the overlapping intervals together and tidy up a bit

```{r}
trt <- pac_ints(trt, trt_start_date, trt_end_date, pat_id, dose_freq, .gap = 1)
trt[, int_grp_id := NULL] # drop the grouping ID from `pac_ints`
setorder(trt, pat_id, trt_start_date) # order for ease of review
trt
class(trt)
```

Then we can split up the intervals to make sure everything is mutually exclusive
```{r}
daily <- trt[dose_freq == "daily"]
weekly <- trt[dose_freq == "weekly"]
monthly <- trt[dose_freq == "monthly"]

weekly
# # remove the parts of the week dosing periods that overlap with the daily dosing intervals
# weekly <- trm_ints(
#   weekly,
#   daily,
#   trt_start_date,
#   trt_end_date,
#   trt_start_date,
#   trt_end_date,
#   pat_id,
#   .gap = 1
#   )
# # remove the parts of the monthly dosing periods that overlap with the daily and weelky dosing intervals
# monthly <- trm_ints(
#   monthly,
#   rbind(daily, weekly),
#   trt_start_date,
#   trt_end_date,
#   trt_start_date,
#   trt_end_date,
#   pat_id,
#   .gap = 1
#   )
# 
# trt2 <- rbindlist(
#   list(
#     trt[dose_freq == "daily"],
#     weekly,
#     monthly
#   )
# )
# setorder(trt2, pat_id, trt_start_date)
# 
# trt2
```
